Integrating brain and biomechanical models - a new paradigm for understanding neuro-muscular control

### 2017-08-19 papers/integration/omm_integ.tex
Current work in progress, downloaded from sharelatex draft

\usepackage{url,hyperref,lineno,microtype}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage[onehalfspacing]{setspace}
\linenumbers

@@

The field of computational neuroscience has provided
many \emph{systems models} of the brain~\cite{several_models}. We
refer to these as \emph{mechanistic computational models}, meaning
models which consist of populations of neural elements, interconnected
in a biologically plausible manner, which simulate the operation of
the brain.

@@

% Short desc of mechanistic computational model
By \emph{mechanistic computational model}, we mean a model which
consists of populations of neural elements, interconnected in a
biologically plausible manner. Such a model may consist of neural
elements that produce a time-dependent membrane voltage and realistic
spiking behaviour (Hodgkin-Huxley, Fizhugh-Nagumo, Izhikevich, Leaky
integrate and fire) in which case the neural element's activity can be
obtained by counting action potentials (or considering their
timing). Alternatively it may (as it does in this work) consist of
rate-coded neural elements, whose activation variable is related to
the spiking rate of a number of neurons (refs).

% Sometimes, ok to just have a brain model
In some cases, it is sufficient to take the activity of an internal
population within the brain model as being representative of the
induced behaviour. For example, a choice made in a \e{go/no-go} task
could be determined from activity in a population within a Basal
Ganglia model
(\mycite{nambu_discharge_1990,kuhn_event-related_2004}). The decision
to \e{go} is selected by a reduction of activity in this population;
maintenance of activity implies \e{no-go}. Error rates in this task
could be compared with experimentally determined error rates in
primate subjects.
% Brain model comparison validates model
If the expression of behaviour by the model is similar when compared
with the experimental data, this serves to validate the model for the
healthy, neuro-typical subject.
% Now describe comparison between model and expt. for diseased
% subject.
If we have confidence in the model's description of the behaviour for
the healthy subject, we can now progress to a comparison between the
model and an unhealthy subject. If we can identify a characteristic
degradation in an aspect of the brain for an unhealthy subject, then
we can seek to modify the model to mirror this degradation. By
comparing the relationship between the model's behaviour as a function
of the degradation parameter with the unhealthy individual's behaviour
as a function of disease progression, we may be able to relate the
disease progression state to the brain parameter and predict the
time-course of the internal degradation of the brain.
% Leave further discussion along these lines to the discussion
% section.

@@

% Motivation here is to study movement disorders by means of a
% comparison between healthy subjects, the healthy model and diseased
% subjects and the diseased model.
Our motivation is to study movement disorders using our brain
models. Because the challenge of integrating brain and biomechanical
models has not, to our knowledge, previously been addressed, we chose
to express some of the simplest, and most studied movements made by
vertebrates: the rapid eye movements known as saccades. Saccades
direct the fovea, the region of the eye with the highest spatial
imaging resolution, towards putative targets of interest. Mammals make
frequent saccades; many thousands every day. These movements are often
described as ballistic: They are commanded by activity in the superior
colliculus which encodes an endpoint for the saccade trajectory
(\mycite{wurtz_superior_1971,hepp_monkey_1993}); once initiated, the
saccade proceeds in a relatively direct and stereotyped (though not
always straight) trajectory (\mycite{van_der_stigchel_eye_2006}) to
the endpoint.
% Why am I mentioning curved saccades? Because it shows that saccades
% can exhibit rich behaviour and that modelling them is worthwhile.
However, numerous studies have shown that saccade trajectories are
affected by distractors (\mycite{mcpeek_competition_2003,
port_sequential_2003, van_der_stigchel_recent_2010}) and may be
modified in-flight to the extent that saccades may be seen to perform
significant, sometimes even U-shaped, curves towards or away from
distractors. This shows that whilst saccades are simple compared to
the movements of limbs, they still provide information about a number
of mechanisms within the brain and an effort to model them accurately
is worthwhile.

% Say what we've done here, in short form.
Parkinson's Disease is a progressive neurological disorder whose main
symptoms relate to motor control. These include resting tremor,
slowness of movements (bradykinesia), difficulty in initiating
movement (akinesia) and rigidity (refs). It is considered to be a
disease of the basal ganglia because during its progression, dopamine
producing neurons in the substantia nigra pars compacta die away
(ref), affecting the the striatum which depends on endogenous dopamine
for correct operation (ref). The basal ganglia plays an important role
in decision making and action selection (ref), including the selection
of saccadic eye movements (\mycite{hikosaka_role_2000}). In the later
stages of Parkinson's Disease, patients show more latent saccades to
targets of interest (ref). They have also been documented as showing
poorer performance in the so-called anti-saccade protocol, in which
the subject is instructed to make a saccade \e{away} from a visual
target. Parkinson's Disease patients make significantly more errors;
they will saccade \e{to} the visual target rather than \e{away} from
it more often than healthy control subjects. To investigate the way in
which degradation in the basal ganglia may be causing these defects we
have developed an integrated brain and biomechanical eye model which
we describe here. The model will initially be used to investigate the
deterioration in anti-saccadic movements which is detectable even in
relatively early-stage Parkinson's Disease.

%See \mycite{Rottach_dynamic_1996} for trajectory information in saccades.

@@

Many neural populations are involved in the coding of saccadic eye
movements, only a very brief overview is given here; for a recent
review, see \mycite{munoz_commentary:_2002}. One pathway takes
information from the retina directly into the superficial layers of
the superior colliculus in the brainstem (\mycite{?}). Activity within
the superior colliculus then excites neurons in the pons, medulla and
rostral midbrain (\mycite{sparks_brainstem_2002}) and then through
channels to the motor neurons which innervate the extraocular muscles
(\mycite{fuchs_firing_1970,sparks_brainstem_2002}). This direct pathway
is responsible for the low latency saccades called express saccades
(\mycite{?}).

% FEF: \mycite{schall_neural_1999}
Information from the retina is also processed by visual cortex which
feeds through to the frontal eye fields in which activity is related
to reflexive and voluntary
saccades \mycite{schall_neural_1999}. Activity build up in the frontal
eye fields is transferred to the intermediate layers of the superior
colliculus
\mycite{?} and is also processed by the basal ganglia, which
participates in the selection of the winning saccade end point
\mycite{?}.

@@

% A very short introduction to the section.
The integrated brain and biomechanical model described here is a
development of the model in \mycite{cope_basal_2017},
referred to here as the \ccg~model. This was a rate-coded neural
network model incorporating retinal populations, frontal eye fields
(FEF), the basal ganglia (BG), and the superior colliculus (SC). In
the \ccg~model, the centroid of the activity in the deep
layers of superior colliculus was assumed to accurately encode the
location of the eye at the end of the
saccade~(\cite{wurtz_activity_1972,robinson_eye_1972,van_gisbergen_collicular_1987,mcilwain_lateral_1982}).
This location was used to recalculate the positions of the luminances in
the eye's frame of reference at each time step. The model included no
brainstem populations other than superior colliculus, nor a
neuromuscular model.

To summarize, the \ccg~model takes as \emph{input} the positions of luminances on a topographic map and produces as output a saccade target.

To the \ccg~model, we have added a rate-coded brainstem model (a ``saccadic burst
generator''), and a biomechanical eye, implemented using the biomechanical modelling framework OpenSim. These will be described below,
but first we will give a description of the modelling framework used.

@@

% Say the \ccg~model was implemented in matlab-brahms & this one is SpineML.
The \ccg~model was originally developed to run on the BRAHMS model
execution framework
(\mycite{mitchinson_brahms:_2010,mitchinson_brahms_2015}). To run a
BRAHMS model, the researcher must develop \e{BRAHMS components} for
the various neural elements. A BRAHMS component is a programmatically
coded implementation of the behaviour of the component. It may have an
arbitrary number of inputs and outputs and may be written in C, C++,
Python or MATLAB. The \ccg~model's components were hand written in C++
and MATLAB. A BRAHMS \e{SystemML} file describes how the different
components connect together and how data is passed between them
(\mycite{mitchinson_brahms:_2010}). The main BRAHMS program first
reads the SystemML file, then dynamically loads all the required
components before executing the system.

In the current work, the \ccg~model was reproduced using the
declarative SpineML markup language (\mycite{richmond_model_2014}),
with the help of the graphical SpineML model editing software
called SpineCreator
(\mycite{cope_spinecreator:_2016}). SpineML,
which is a development of the NineML specification
(\mycite{incf_task_force_on_multi-scale_modeling_network_2011}),
describes neural populations and their projections in a highly
structured format in which neuron bodies, pre- and post-synapses are
described in terms of \e{SpineML components}. These are similar to the
components provided by BRAHMS, but in this case, the components are an
XML description of the functionality of the component, rather than a
programmatic implementation, with one XML file per component. A
SpineML \e{network layer} file then describes which components are
used in the model, and how they are connected together. Finally, a
number of SpineML \e{experiment layer} files specify how the model
described in the network layer can be executed. In the experiment
layer, the execution duration and timestep can be specified, along
with input conditions, connection lesions and component parameter
updates. A complete description of SpineML is given
in \mycite{richmond_model_2014}.

@@

As a declarative format for model specification, SpineML is agnostic
about how the model is executed. A number of simulation
engines can be utilised, including DAMSON (\mycite{richmond_damson_2015}), GeNN
(\mycite{nowotny_flexible_2011,nowotny_spineml_2014}) and BRAHMS (used
here). The simulation engine incorporating BRAHMS is called \stob~
(\mycite{cope_spineml_2_brahms_2015-1}). \stob~is a collection of XSLT
stylesheets which first generate and compile C++ BRAHMS components
from the SpineML component layer description files. \stob~then uses
the SpineML network and experiment layer files to generate a BRAHMS
SystemML description of the model. Finally, \stob~executes the model,
now described entirely as a BRAHMS system, via a call to the BRAHMS
binary. A number of additional hand-written components are present
in \stob~providing the inputs (constant inputs, time-varying inputs,
etc) which the modeller specifies in the experiment layer.

In addition to the brain model components, all of which are
code-generated using \stob~as described above, two hand-written
components are integrated into the model: The biomechanical eye model
and a sensory input component. The sensory input component takes the
eye's rotational state and the state of the experimental luminances
and projects a retinotopic activity map into the brain model. Both of
these BRAHMS components were hand-written in C++. To incorporate these
components into the SpineML model, a model-specific modification was
applied in \stob~(\mycite{cope_spineml_2_brahms_2015}) making use of
a \stob~feature designed for this purpose (The external.xsl file).

@@

The brain model, excluding the brainstem, is a re-implementation of
the \ccg~model. Figure \ref{fig:brain_model} shows the layout of the
populations and the interconnections between them. Each population of
2500 elements is arranged as a 50 by 50 grid which forms a map in a
retinotopic co-ordinate system, roughly matching the known layout of
the superior colliculus~(\mycite{robinson_eye_1972}). % FIXME: and frontal eye field?

@@

The LINlinear component governs the activation $a$ with a first order
leaky integrator differential equation:
\[
   \dot{a} = {1 \over \tau}(a_{in}-a)
\]
where $\tau$ is the time constant for the neural activation and
$a_{in}$ is the input to the neural element. $a_{in}$ is defined by an
activation input and a shunting inhibition input according to:
\[
   a_{in} = A(1-s_a)+\alpha R_N
\]
Here, $A$ is the activation input and $s_a$ is the shunting inhibition
state variable whose value is related to the shunting input, $S$ by
\[
   s_a = \begin{cases}
      S & S\leq 1 \\
      1 & S > 1
   \end{cases}
\]
$R_N$ is a random number drawn from a standard normal distribution
($\sigma$=1, $\mu$=0) and introduces noise to the activation of the
neural element, with the parameter $\alpha$ controlling the noise
amplitude.

The output, $y$, of LINlinear is related to the activation $a$ by the
piecewise linear transfer function
\[
   y(a) = \begin{cases}
      0   & a < c \\
      a-c & c \leq a \leq 1+c \\
      1   & a > 1+c
   \end{cases}
\]
where $c$ is a parameter defining the slope of the transfer
function. At this point, the naming scheme for the component becomes
apparent; this is a Leaky Integrator with a linear transfer function.

The LINret component used for the retinal populations is again similar
to the LINlinear component, but with no intrinsic noise and no
shunting inhibitory input. It has a neural input which is identical to
the activation input $A$:
\[
   a_{in} = A
\]
The LINexp component is a leaky integrator with an exponential
transfer function. It shares the same differential equation with
LINlinear, but has a different input equation and a different output
transfer function. It has the following equation for the neural
element input $a_{in}$:
\[
   a_{in} = [A+N(a-V_{r}^{-})] (1-S) + 0.01 R_N
\]
where $A$ is the activation input and $N$ is an input which is
modulated by $V_{r}^{-}$, a reversal potential, and $a$, the current
activation of the element. These inputs are summed and then reduced by
a factor which is dependent on $S$, the shunting input. As in
LINlinear, $R_N$ introduces normally distributed noise to the element.

The output, $y$, of the LINexp component is given by
\[
   y(a) = \begin{cases}
      \mathrm{e}^{a}-0.9   & \mathrm{e}^a \leq 1+0.9 \\
      1   & \mathrm{e}^a > 1+0.9
   \end{cases}
\]
This component is used in the STN population, as it gives a more
physiologically accurate f-I
behaviour \mycite{wilson_model_2004,bevan_mechanisms_1999,hallworth_apamin-sensitive_2003}
which has been shown to allow the mapping of the basal ganglia network
architecture onto an optimal decision making
model \mycite{bogacz_basal_2007}.

The D1MSN and D2MSN components are both leaky integrators, similar to
LINlinear. They differ in that they have no shunting inhibition and a
dopamine parameter that modulates the input activation, so
that their equations for $a_{in}$ are thus:
\[
   a_{in}^{D1} = (0.2 + d)A + 0.01 R_N
\]
\[
   a_{in}^{D2} = (1 - d)A + 0.01 R_N
\]
where $d$ is a dopamine parameter. Varying dopamine from 0 to 1
enhances the activation in the D1 model, whereas it decreases
the activation of the D2 model elements, in line with experimental
observations~\cite{refs}. Note that the equation for $a_{in}^{D1}$
differs from that used in the \ccg~model, for which the
cortico-striatal weights are multiplied by $(1+d)$ rather than
$(0.2+d)$.

The equations given above are applied to each element in a
population. The value of the activation $A$ (and where relevant, the
shunting input, $S$) is determined by summing the weighted
inputs to the population:
\[
A = \sum_{i}w_i^{act} x_i^{act}
\]
\[
S = \sum_{i}w_i^{sh} x_i^{sh}
\]

@@

Each population of 2500 neural elements is arranged in a 50 by 50
grid, with positions on the grid representing a retinotopic
mapping. Locations on the grid specify radial and azimuthal angles $r$
and $\phi$ for the location of activity on the retina relative to the
fovea. $r$ and $\phi$ can be converted into rotations about the $x$ and
$y$ Cartesian axes ($\theta_x$ and $\theta_y$). The result of a
positive rotation about the $x$ axis in a right-handed Cartesian
coordinate system results in an upward movement of the fovea's view
of the environment. Note that a positive rotation about the $y$ axis
results in a \emph{leftward} movement of the eye's focal point, with
respect to the world.

Note that the mapping encompases the entire visual field. The value of
$\phi$ is allowed to vary from 0$\deg$ to 360$\deg$ along its axis.
Effectively, the two contralateral colliculi are incorporated into a
single, square map, avoiding the need to carry out `colliculus
gluing'~\cite{tabareau_geometry_2007}.

The magnification factor, $M_f$, gives the relationship between the
radial angle $r$ and the retinal neural density. It is given by:
\[ \label{eq:mf}
   M_f = {W_{nfs} \over E_2 ln ({W_{fov} \over 2 E_2} + 1)}
\]
where $W_{nfs}$ is the width of the retinotopic neural fields (50),
$W_{fov}$ is the eye's field of view, which is set to 61 degrees, a
reduction from the biophysically accurate 150 degrees due to the small
number of neurons in the retinotopic neural field. $E_2$ is the
eccentricity (in degrees) at which the retinal neural density has
halved compared with the foveal region.

With $M_f$ calculated, the mapping from the retinotopic radial and
angular components, $r$ and $\phi$ to $\theta_x$ and $\theta_y$ is:
\[
   \theta_x = E_2 (e^{{r \over M_f E_2}} - 1).\cos({2 \pi \phi \over
   W_{nfs}}) \tag{1}
\]
\[
   \theta_y = E_2 (e^{{r \over M_f E_2}} - 1).\sin({2 \pi \phi \over W_{nfs}}) \tag{2}
\]

It is straightforward to show that the reverse mapping is given by:
\[
   \phi = {W_{nfs} \over 2 \pi}\arctan({\theta_y\over\theta_x}) \tag{3}
\]
\[
   r = M_f E_2 \ln ({1\over E_2}\sqrt{{\theta_x}^2 + {\theta_y}^2}+1) \tag{4}
\]

Figure \ref{fig:mapping} shows the result of the mapping for a view of
two cross-shaped luminances. One cross illuminates the fovea, which
results in a large comb-shape of activity. The more peripheral cross
produces (in FEF) an indistinct object centered at a larger value of
r.

>>>>

\usepackage{url,hyperref,lineno,microtype}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage[onehalfspacing]{setspace}
\usepackage{gensymb} % For \degree

\linenumbers

@@

The field of computational neuroscience has provided
many \e{systems models} of the brain~\cite{arai_two-dimensional_1994,gancarz_neural_1998,hazy_towards_2007,blenkinsop_frequency_2017}. We
refer to these as \e{mechanistic computational models}, meaning
models which consist of populations of neural elements, interconnected
in a biologically plausible manner, which simulate the operation of
the brain.

@@

% Short desc of mechanistic computational model
By \e{mechanistic computational model}, we mean a model which
consists of populations of neural elements, interconnected in a
biologically plausible manner. Such a model may consist of neural
elements that produce a time-dependent membrane voltage and realistic
spiking behaviour (Hodgkin-Huxley, Fizhugh-Nagumo, Izhikevich, Leaky
integrate and fire) in which case the neural element's activity can be
obtained by counting action potentials (or considering their
timing). Alternatively it may (as it does in this work) consist of
rate-coded neural elements, whose activation variable is related to
the spiking rate of a number of neurons (refs).

% Sometimes, ok to just have a brain model
In some cases, it is sufficient to take the activity of an internal
population within the brain model as being representative of the
induced behaviour. For example, a choice made in a \e{go/no-go} task
could be determined from activity in a population within a Basal
Ganglia model
(\cite{nambu_discharge_1990,kuhn_event-related_2004}). The decision
to \e{go} is selected by a reduction of activity in this population;
maintenance of activity implies \e{no-go}. Error rates in this task
could be compared with experimentally determined error rates in
primate subjects.
% Brain model comparison validates model
If the expression of behaviour by the model is similar when compared
with the experimental data, this serves to validate the model for the
healthy, neuro-typical subject.
% Now describe comparison between model and expt. for diseased
% subject.
If we have confidence in the model's description of the behaviour for
the healthy subject, we can now progress to a comparison between the
model and an unhealthy subject. If we can identify a characteristic
degradation in an aspect of the brain for an unhealthy subject, then
we can seek to modify the model to mirror this degradation. By
comparing the relationship between the model's behaviour as a function
of the degradation parameter with the unhealthy individual's behaviour
as a function of disease progression, we may be able to relate the
disease progression state to the brain parameter and predict the
time-course of the internal degradation of the brain.
% Leave further discussion along these lines to the discussion
% section.

@@

% Motivation here is to study movement disorders by means of a
% comparison between healthy subjects, the healthy model and diseased
% subjects and the diseased model.
Our motivation is to study movement disorders using our brain
models. Because the challenge of integrating brain and biomechanical
models has not, to our knowledge, previously been addressed, we chose
to express some of the simplest, and most studied movements made by
vertebrates: the rapid eye movements known as saccades. Saccades
direct the fovea, the region of the eye with the highest spatial
imaging resolution, towards putative targets of interest. Mammals make
frequent saccades; many thousands every day. These movements are often
described as ballistic: They are commanded by activity in the superior
colliculus which encodes an endpoint for the saccade trajectory
(\cite{wurtz_superior_1971,hepp_monkey_1993}); once initiated, the
saccade proceeds in a relatively direct and stereotyped (though not
always straight) trajectory (\cite{van_der_stigchel_eye_2006}) to
the endpoint.
% Why am I mentioning curved saccades? Because it shows that saccades
% can exhibit rich behaviour and that modelling them is worthwhile.
However, numerous studies have shown that saccade trajectories are
affected by distractors (\cite{mcpeek_competition_2003,
port_sequential_2003, van_der_stigchel_recent_2010}) and may be
modified in-flight to the extent that saccades may be seen to perform
significant, sometimes even U-shaped, curves towards or away from
distractors. This shows that whilst saccades are simple compared to
the movements of limbs, they still provide information about a number
of mechanisms within the brain and an effort to model them accurately
is worthwhile.

% Say what we've done here, in short form.
Parkinson's Disease is a progressive neurological disorder whose main
symptoms relate to motor control. These include resting tremor,
slowness of movements (bradykinesia), difficulty in initiating
movement (akinesia) and rigidity (refs). It is considered to be a
disease of the basal ganglia because during its progression, dopamine
producing neurons in the substantia nigra pars compacta die away
(ref), affecting the the striatum which depends on endogenous dopamine
for correct operation (ref). The basal ganglia plays an important role
in decision making and action selection (ref), including the selection
of saccadic eye movements (\cite{hikosaka_role_2000}). In the later
stages of Parkinson's Disease, patients show more latent saccades to
targets of interest (ref). They have also been documented as showing
poorer performance in the so-called anti-saccade protocol, in which
the subject is instructed to make a saccade \e{away} from a visual
target. Parkinson's Disease patients make significantly more errors;
they will saccade \e{to} the visual target rather than \e{away} from
it more often than healthy control subjects. To investigate the way in
which degradation in the basal ganglia may be causing these defects we
have developed an integrated brain and biomechanical eye model which
we describe here. The model will initially be used to investigate the
deterioration in anti-saccadic movements which is detectable even in
relatively early-stage Parkinson's Disease.

%See \cite{Rottach_dynamic_1996} for trajectory information in saccades.

@@

Many neural populations are involved in the coding of saccadic eye
movements, only a very brief overview is given here; for a recent
review, see \cite{munoz_commentary:_2002}. One pathway takes
information from the retina directly into the superficial layers of
the superior colliculus in the brainstem (\cite{sterling_receptive_1971,linden_massive_1983,wu_involvement_1994}).
Activity within
the superior colliculus then excites neurons in the pons, medulla and
rostral mid-brain (\cite{sparks_brainstem_2002}) and then through
channels to the motor neurons which innervate the extraocular muscles
(\cite{fuchs_firing_1970,sparks_brainstem_2002}). This direct pathway
is responsible for the low latency saccades called express saccades
(\cite{schiller_effect_1987,edelman_activity_1996}).

% FEF: \cite{schall_neural_1999}
Information from the retina is also processed by visual cortex which
feeds through to the frontal eye fields in which activity is related
to reflexive and voluntary
saccades \cite{schall_neural_1999}. Activity build up in the frontal
eye fields is transferred to the intermediate layers of the superior
colliculus
\cite{stanton_frontal_1988-1} and is also processed by the basal ganglia, which
participates in the selection of the winning saccade end point
\cite{stanton_frontal_1988}.

@@

% A very short introduction to the section.
The integrated brain and biomechanical model described here is a
development of the model in \cite{cope_basal_2017},
referred to here as the \ccg~model. This was a rate-coded neural
network model incorporating retinal populations, frontal eye fields
(FEF), the basal ganglia (BG), and the superior colliculus (SC). In
the \ccg~model, the centroid of the activity in the deep
layers of superior colliculus was assumed to accurately encode the
location of the eye at the end of the
saccade~(\cite{wurtz_activity_1972,robinson_eye_1972,van_gisbergen_collicular_1987,mcilwain_lateral_1982}).
This location was used to recalculate the positions of the luminances in
the eye's frame of reference at each time step. The model included no
brainstem populations other than superior colliculus, nor a
neuromuscular model.

To summarize, the \ccg~model takes as \e{input} the positions
of luminances on a topographic map and produces as output a saccade 
target.

To the \ccg~model, we have added a rate-coded brainstem model (a ``saccadic
burst generator''), and a biomechanical eye, implemented using the
biomechanical modelling framework OpenSim. These will be described below,
but first we will give a description of the co-ordinate systems and the
modelling framework used.

\subsubsection{Coordinates in the world}

Before describing the biomechanical eye and the brain model, which consisted
of retinotopically mapped neural sheets, we describe the coordinate system
used in the world. The eye was located at the origin of a 3 dimensional, right-handed 
Cartesian co-ordinate system, with its fovea directed in the $-z$ direction. 
There was a notional spherical screen which was also centered at the origin of the
co-ordinate system and had a radius of 50 (in arbitrary units). The \e{fixation 
point} was the point on the
screen at which the eye was initially directed.
Onto the screen were projected luminances, each of which having a position 
described by two co-ordinates; $\theta_x$, a 
rotation of the horizon plane about the $x$ axis, and $\theta_y$, a rotation 
of the meridian plane about the $y$ axis. The position is the intersection
of these rotated planes with the spherical screen (disregarding
the intersection point of these three surfaces behind the eye).
%
Note that a luminance with positive $\theta_x$ was above the horizon of this world; 
one whose $\theta_y$ was positive lay to the left of the world's meridian.

Luminances were crosses of height and width subtending 6$\degree$ and whose
`bars' were 2$\degree$ thick. Luminances were oriented like $+$ symbols with
their vertical bar aligned with the meridian plane and their horizontal bar 
aligned with the horizon plane described above.

The eye's frame of reference was initially aligned with the world's frame
of reference. 
At each timestep, the eye's rotational state ($R_x$, $R_y$, $R_z$) was 
used to translate the three dimensional
Cartesian co-ordinates of the luminances in the world frame into co-ordinates
in the eye frame. The luminance coordinates in the eye's frame of reference
were used to determine the input to the brain model.

@@

% Say the \ccg~model was implemented in matlab-brahms & this one is SpineML.
The \ccg~model was originally developed to run on the BRAHMS model
execution framework
(\cite{mitchinson_brahms:_2010,mitchinson_brahms_2015}). To run a
BRAHMS model, the researcher must develop \e{BRAHMS components} for
the various neural elements. A BRAHMS component is a programmatically
coded implementation of the behaviour of the component. It may have an
arbitrary number of inputs and outputs and may be written in C, C++,
Python or MATLAB. The \ccg~model's components were hand written in C++
and MATLAB. A BRAHMS \e{SystemML} file describes how the different
components connect together and how data is passed between them
(\cite{mitchinson_brahms:_2010}). The main BRAHMS program first
reads the SystemML file, then dynamically loads all the required
components before executing the system.

In the current work, the \ccg~model was reproduced using the
declarative SpineML markup language (\cite{richmond_model_2014}),
with the help of the graphical SpineML model editing software
called SpineCreator
(\cite{cope_spinecreator:_2016}). SpineML,
which is a development of the NineML specification
(\cite{incf_task_force_on_multi-scale_modeling_network_2011}),
describes neural populations and their projections in a highly
structured format in which neuron bodies, pre- and post-synapses are
described in terms of \e{SpineML components}. These are similar to the
components provided by BRAHMS, but in this case, the components are an
XML description of the functionality of the component, rather than a
programmatic implementation, with one XML file per component. A
SpineML \e{network layer} file then describes which components are
used in the model, and how they are connected together. Finally, a
number of SpineML \e{experiment layer} files specify how the model
described in the network layer can be executed. In the experiment
layer, the execution duration and timestep can be specified, along
with input conditions, connection lesions and component parameter
updates. A complete description of SpineML is given
in \cite{richmond_model_2014}.

@@

As a declarative format for model specification, SpineML is agnostic
about how the model is executed. A number of simulation
engines can be utilised, including DAMSON (\cite{richmond_damson_2015}), GeNN
(\cite{nowotny_flexible_2011,nowotny_spineml_2014}) and BRAHMS (used
here). The simulation engine incorporating BRAHMS is called \stob~
(\cite{cope_spineml_2_brahms_2015-1}). \stob~is a collection of XSLT
stylesheets which first generate and compile C++ BRAHMS components
from the SpineML component layer description files. \stob~then uses
the SpineML network and experiment layer files to generate a BRAHMS
SystemML description of the model. Finally, \stob~executes the model,
now described entirely as a BRAHMS system, via a call to the BRAHMS
binary. A number of additional hand-written components are present
in \stob~providing the inputs (constant inputs, time-varying inputs,
etc) which the modeller specifies in the experiment layer.

In addition to the brain model components, all of which are
code-generated using \stob~as described above, two hand-written
components are integrated into the model: The biomechanical eye model
and a sensory input component. The sensory input component takes the
eye's rotational state and the state of the experimental luminances
and projects a retinotopic activity map into the brain model. Both of
these BRAHMS components were hand-written in C++. To incorporate these
components into the SpineML model, a model-specific modification was
applied in \stob~(\cite{cope_spineml_2_brahms_2015}) making use of
a \stob~feature designed for this purpose (The external.xsl file).

@@

The brain model, excluding the brainstem, is a re-implementation of
the \ccg~model. Figure \ref{fig:brain_model} shows the layout of the
populations and the interconnections between them. Each population of
2500 elements is arranged as a 50 by 50 grid which forms a map in a
retinotopic co-ordinate system, roughly matching the known layout of
the superior colliculus~(\cite{robinson_eye_1972}). % FIXME: and frontal eye field?

@@

The LINlinear component governs the activation $a$ with a first order
leaky integrator differential equation:
\begin{equation}
   \dot{a} = \frac{1}{\tau}(a_{in}-a)
\end{equation}
where $\tau$ is the time constant for the neural activation and
$a_{in}$ is the input to the neural element. $a_{in}$ is defined by an
activation input and a shunting inhibition input according to:
\begin{equation}
   a_{in} = A(1-s_a)+\alpha R_N
\end{equation}
Here, $A$ is the activation input and $s_a$ is the shunting inhibition
state variable whose value is related to the shunting input, $S$ by
\begin{equation}
   s_a = \begin{cases}
      S & S\leq 1 \\
      1 & S > 1
   \end{cases}
\end{equation}
$R_N$ is a random number drawn from a standard normal distribution
($\sigma$=1, $\mu$=0) and introduces noise to the activation of the
neural element, with the parameter $\alpha$ controlling the noise
amplitude.

The output, $y$, of LINlinear is related to the activation $a$ by the
piecewise linear transfer function
\begin{equation}
   y(a) = \begin{cases}
      0   & a < c \\
      a-c & c \leq a \leq 1+c \\
      1   & a > 1+c
   \end{cases}
\end{equation}
where $c$ is a parameter defining the slope of the transfer
function. At this point, the naming scheme for the component becomes
apparent; this is a Leaky Integrator with a linear transfer function.

The LINret component used for the retinal populations is again similar
to the LINlinear component, but with no intrinsic noise and no
shunting inhibitory input. It has a neural input which is identical to
the activation input $A$:
\begin{equation}
   a_{in} = A
\end{equation}
The LINexp component is a leaky integrator with an exponential
transfer function. It shares the same differential equation with
LINlinear, but has a different input equation and a different output
transfer function. It has the following equation for the neural
element input $a_{in}$:
\begin{equation}
   a_{in} = [A+N(a-V_{r}^{-})] (1-S) + 0.01 R_N
\end{equation}
where $A$ is the activation input and $N$ is an input which is
modulated by $V_{r}^{-}$, a reversal potential, and $a$, the current
activation of the element. These inputs are summed and then reduced by
a factor which is dependent on $S$, the shunting input. As in
LINlinear, $R_N$ introduces normally distributed noise to the element.

The output, $y$, of the LINexp component is given by
\begin{equation}
   y(a) = \begin{cases}
      \mathrm{e}^{a}-0.9   & \mathrm{e}^a \leq 1+0.9 \\
      1   & \mathrm{e}^a > 1+0.9
   \end{cases}
\end{equation}
This component is used in the STN population, as it gives a more
physiologically accurate f-I
behaviour \cite{wilson_model_2004,bevan_mechanisms_1999,hallworth_apamin-sensitive_2003}
which has been shown to allow the mapping of the basal ganglia network
architecture onto an optimal decision making
model \cite{bogacz_basal_2007}.

The D1MSN and D2MSN components are both leaky integrators, similar to
LINlinear. They differ in that they have no shunting inhibition and a
dopamine parameter that modulates the input activation, so
that their equations for $a_{in}$ are thus:
\begin{equation}
   a_{in}^{D1} = (0.2 + d)A + 0.01 R_N
\end{equation}
\begin{equation}
   a_{in}^{D2} = (1 - d)A + 0.01 R_N
\end{equation}
where $d$ is a dopamine parameter. Varying dopamine from 0 to 1
enhances the activation in the D1 model, whereas it decreases
the activation of the D2 model elements, in line with experimental
observations~\cite{refs}. Note that the equation for $a_{in}^{D1}$
differs from that used in the \ccg~model, for which the
cortico-striatal weights are multiplied by $(1+d)$ rather than
$(0.2+d)$.

The equations given above are applied to each element in a
population. The value of the activation $A$ (and where relevant, the
shunting input, $S$) is determined by summing the weighted
inputs to the population:
\begin{equation}
A = \sum_{i}w_i^{act} x_i^{act}
\end{equation}
\begin{equation}
S = \sum_{i}w_i^{sh} x_i^{sh}
\end{equation}

@@

Each population of 2500 neural elements was arranged in a 50 by 50
grid, with positions on the grid representing a retinotopic
mapping similar to that found empirically both in the superior 
colliculus (\cite{ottes_visuomotor_1986}) and in visual cortex (\cite{eric_l._schwartz_computational_1980}) and assumed in
this work to persist throughout the oculomotor system.

In a retinotopic mapping, the Cartesian coordinates of the light-sensitive
cells in the retina, whose density varies with distance from the fovea,
are transformed into the Cartesian co-ordinates of 
the correspondingly active cells on the colliculus. The mapping ensures 
that an even density of cells can be maintained in the colliculus, but 
ensures that a single group of proximal, active, retinal neurons will 
always activate a single, proximal group of neurons on the collicular 
surface.

The mapping turns out to resemble polar coordinates. That is, one 
axis of the collicular surface specifies the eccentricity of
a retinal location (how far it is from the fovea) and the second axis
specifies the rotational angle of the retinal location; we therefore
use the convention 
of referring to the eccentricity axis on the colliculus as $r$ and the 
rotation axis as $\phi$.

The \e{cortical magnification factor}, $M(r)$, gives the relationship
between the
radial eccentricity $r$ and the retinal neural density. As in 
\cite{cope_basal_2017}, we use a first-order approximation of the form
for $M(r)$ given in  \cite{rovamo_estimation_1979}:
\begin{equation} \label{eq:cmf}
M(r) = \frac{M_f}{1+\frac{r}{E_2}}
\end{equation}
The foveal magnification, $M_f$, is the magnification of the most central
region of the retina and has a value in the human of about
7.8 mm/$\degree$~\cite{rovamo_estimation_1979}.

In our model, $M_f$ is related to $W_{nfs}$, the width of the retinotopic neural 
field, $W_{fov}$, the width of the eye's field of view and $E_2$, the 
eccentricity at which the retinal density has halved by:
\begin{equation} \label{eq:fm}
   M_f = \frac {W_{nfs}} {E_2\;\ln\left(\frac{W_{fov}}{2 E_2} + 1\right)}
\end{equation}
Here, $W_{nfs}$ is 50, $W_{fov}$ is set to 61$\degree$, a
reduction from the biophysically accurate 150$\degree$ due to the small
number of neurons in the retinotopic neural field. $E_2$ is
2.5~\cite{cope_basal_2017,slotnick_electrophysiological_2001}.
